---
title: "Lab 03 - Nobel laureates"
author: "Lindley Slipetz"
date: "02/09/2021"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE, warning = FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

```{r glimpse}
glimpse(nobel)
```
There are 935 rows and 26 columns. The rows represent nobel laureates.

### Exercise 2

```{r noble_living}
nobel_living <- nobel %>%
  filter(is.na(country) == FALSE,
         gender != "org",
         is.na(died_date))

```
### Exercise 3

```{r create_cu}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", 
                         "Other")
                )
```

```{r filter_field}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

```{r bar_graph}
ggplot(nobel_living, aes(x=factor(country_us)))+
  geom_bar(stat="count", width=0.7, fill="black")+
  facet_wrap(~category, nrow = 2) +
  labs(x = "Country of Nobel", 
      title = "US vs other countries for Nobel laureates by category") +
  coord_flip()
```
The claim that most nobels come from the US is supported. It remains to be seen if, of those from the US, they come from other countries.

### Exercise 4

```{r create_bcu}
nobel_living <- nobel_living %>%
  mutate(
    born_country_us = if_else(born_country == "USA",
                              "USA", "Other") 
                )
count(nobel_living, vars = born_country_us)
```

123 were born in countries other than the US, while 105 were born in the US.

### Exercise 5

```{r bar_graph_2}
ggplot(nobel_living, aes(x=factor(country_us), fill = born_country_us))+
  geom_bar(stat="count", width=0.7)+
  facet_wrap(~category, nrow = 2) +
  labs(x = "Country of Nobel", fill = "Birth place",
      title = "US vs other countries for Nobel laureates by category") +
  coord_flip() +
    scale_fill_manual(values=c("black", "#999999"))
```
Except for Economics, each category has a sizeable proportion of nobel laureates born outside the US, even if they received the prize in the US. It seems that if we added the black portions of the bar plots together, it would be greater than the portions with US-born nobel laureates. Hence, I agree with the Buzzfeed article.

### Exercise 6

```{r where_born}
nobel %>%
  filter(born_country != "USA", 
         country == "USA") %>%
  count(vars = born_country) %>%
  arrange(desc(n))
```

The most common country is the UK.
